---
title: "UKB mtDNA-CN Analysis"
output:
  html_document:
    code_folding: "hide"
  df_print: paged
---



```{r setup, include=FALSE} 
library(tidyverse)
library(magrittr)
library(data.table)
library(pander)
library(plotly)
library(KRIS)
library(MASS)
library(sfsmisc)
library(DT)
library(ggpubr)
library(ggstatsplot)
select <- dplyr::select

# formatting functions
sensible_names <- function(x){
  colnames(x) %<>%
    str_replace_all('[ -//]+', '.') %>%
    str_to_lower()
  x
}

knitr::opts_chunk$set(tidy="styler")
```

### Check Batch effects

```{r,message=FALSE, warning=FALSE, eval=TRUE}
#WGS cohort of approximately 200k individuals (released on late 2021)
ukbdata <- fread("/vast/projects/bahlo_ukbiobank/app36610_HDA/mitochondria/20231124_mtDNA_ukbb_summary.csv")
ukbdata <- ukbdata %>% 
  filter(Parkinsons=="Y")  %>% 
  mutate(Data="Batch1")  %>% 
  dplyr::select(sample,Data,genetic_sex, mt,Blood_Collection_Age,logmt)
#new ~500k release (released on late 2023) additional PD cases data 
additional_pd_case1 <- fread("/vast/projects/bahlo_ukbiobank/app36610_HDA/mitochondria/20240220_mtDNA_ukbb_additional_pd_summary.csv")
additional_pd_case1 <- additional_pd_case1 %>% 
  mutate(Data="Batch2") %>% 
  dplyr::select(sample,Data,genetic_sex, mt,Blood_Collection_Age,logmt)
additional_pd_case2 <- fread("/vast/projects/bahlo_ukbiobank/app36610_HDA/mitochondria/20240228_mtDNA_ukbb_additional_pd_summary.csv")
additional_pd_case2 <- additional_pd_case2 %>% 
  mutate(Data="Batch2") %>% 
  dplyr::select(sample,Data,genetic_sex, mt,Blood_Collection_Age,logmt)
additional_pd_case <- do.call(rbind,list(additional_pd_case1,additional_pd_case2))
all <- do.call(rbind,list(ukbdata,additional_pd_case))
all %>% 
  dplyr::select(sample,Data,genetic_sex, mt,Blood_Collection_Age,logmt) %>% 
  pivot_longer(cols = mt:logmt,
               names_to = "Type",
               values_to = "Value") %>% 
  filter(Type=="logmt") %>% 
  ggplot(aes(x=Value,fill=Data))+
  geom_density()+
  labs(x="logmt")

all %>% 
  dplyr::select(sample,Data,genetic_sex, mt,Blood_Collection_Age,logmt) %>% 
  pivot_longer(cols = mt:logmt,
               names_to = "Type",
               values_to = "Value") %>% 
  filter(Type=="logmt") %>% 
  ggplot(aes(x=Data,y=Value,fill=Data))+
  geom_boxplot()+
  labs(y="logmt")

old <- all %>% 
  filter(Data=="Batch1")
new <- all %>% 
  filter(Data=="Batch2")
t.test(old$logmt, y = new$logmt,
       alternative = "two.sided")
```


<br>

### Check log(mtDNA-CN) distribution

<span style="color:blue">**UKBBdata table (combine boold cell composition & remove NA)**</span>

```{r,message=FALSE, warning=FALSE, eval=TRUE}
# load mtDNA-CN results
ukbdata <- fread("/vast/projects/bahlo_ukbiobank/app36610_HDA/mitochondria/20231124_mtDNA_ukbb_summary.csv")
ukbdata <- ukbdata %>% 
  mutate(Batch=1) %>% 
  dplyr::select(sample,mt,genetic_sex,Blood_Collection_Age,logmt,Batch,PC1,PC2,PC3,PC4,PC5,Parkinsons)

new1 <- fread("/vast/projects/bahlo_ukbiobank/app36610_HDA/mitochondria/20240220_mtDNA_ukbb_additional_pd_summary.csv")
new1 <- new1 %>% 
  mutate(Batch=2) %>% 
  dplyr::select(sample,mt,genetic_sex,Blood_Collection_Age,logmt,Batch,PC1,PC2,PC3,PC4,PC5,Parkinsons)
new2 <- fread("/vast/projects/bahlo_ukbiobank/app36610_HDA/mitochondria/20240228_mtDNA_ukbb_additional_pd_summary.csv")
new2 <- new2 %>% 
  mutate(Batch=2) %>% 
  dplyr::select(sample,mt,genetic_sex,Blood_Collection_Age,logmt,Batch,PC1,PC2,PC3,PC4,PC5,Parkinsons)
ukbdata <- do.call(rbind,list(ukbdata,new1,new2))
ukbdata <- ukbdata %>% 
  dplyr::select(sample,mt,genetic_sex,Blood_Collection_Age,logmt,Batch,PC1,PC2,PC3,PC4,PC5,Parkinsons) %>% 
  mutate(PD_diagnosis = ifelse(Parkinsons=="Y","Case","Control")) %>% 
  dplyr::select(-Parkinsons) %>% 
  na.omit()
colnames(ukbdata) <- c("sample", "mt", "sex", "age","logmt", "Batch", "PC1","PC2", "PC3","PC4", "PC5", "diagnosis")
#boold cell composition
data <- fread("/vast/projects/bahlo_ukbiobank/app36610_HDA/extracts/20240118_673741_ukb_blood.csv.gz")
data <- data %>% 
  dplyr::select(eid,matches("0-0.0")) %>% 
  dplyr::select(eid,where(is.numeric))
data <- data %>%   
  pivot_longer(!eid,
               names_to = "FieldIDNew",
               values_to = "Value")
dic <- fread("/vast/projects/bahlo_ukbiobank/app36610_HDA/extracts/20240219_Data_Dictionary_Showcase.csv")
dic <- dic %>% 
  dplyr::select(FieldID,Field)
dic <- dic %>% 
  mutate(FieldIDNew=paste0(FieldID,"-0.0"))
data <- data %>% 
  left_join(dic)
data <- data %>% 
  dplyr::select(eid,Field,Value) %>% 
  pivot_wider(names_from = Field,
              values_from = Value)
ukbdata <- ukbdata %>% 
  left_join(data,by=c("sample"="eid"))
ukbdata %>% 
   as.tibble()
mtDNA_ukbb_summary_PD_blood <- ukbdata
ukbdata <- mtDNA_ukbb_summary_PD_blood
range(ukbdata$logmt)
hist(ukbdata$logmt)

```

<br>

<hr>

### Boold cell composition QC

<span style="color:blue">**UKBB boold cell measurements**</span>

```{r, message=FALSE, warning=FALSE, eval=TRUE}
blood_measure <- read.csv("Blood_Measurements.csv")
blood_measure$UKBBLink <- paste0('<a href="',blood_measure$UKBBLink,'">', 'UKBB_', blood_measure$FieldID ,"</a>")
datatable(blood_measure,
          extensions = 'Buttons',
          options = list(pageLength = 10,
                          dom = 'Blfrtip',
                          buttons = c('copy', 'csv', 'excel')),
          escape = F)
```

<br>

<span style="color:blue">Remove high missing blood cell type > 1%</span>

```{r, message=FALSE, warning=FALSE, eval=TRUE}
#Remove high missing cell type
data <- mtDNA_ukbb_summary_PD_blood
data_miss <- data.frame(miss=colSums(is.na(data))/nrow(data)*100) %>% 
  filter(miss>1)
data_miss
data <- data %>% 
  dplyr::select(-rownames(data_miss))
```

<span style="color:blue">Select cell types percentage with proportion > 1%</span>

```{r, message=FALSE, warning=FALSE, eval=TRUE}
cell_percentage <- data %>% 
  dplyr::select(sample,matches("percentage"))
vars <- c(colnames(cell_percentage[,2:8])[colMeans(cell_percentage[,2:8], na.rm = TRUE) < 1])
vars
data <- data %>% 
  dplyr::select(-c(vars,`Basophill count`,`Nucleated red blood cell count`))
```



<span style="color:blue">Remove all the RBC measurements</span>

The blood measurements associated with Red Blood Cells (RBCs) are:
Red blood cell (erythrocyte) count
Haemoglobin concentration
Haematocrit percentage
Mean corpuscular volume
Mean corpuscular haemoglobin
Mean corpuscular haemoglobin concentration
Red blood cell (erythrocyte) distribution width
Nucleated red blood cell count
Nucleated red blood cell percentage
Mean reticulocyte volume
Mean sphered cell volume
These measurements provide valuable information about the quantity, size, and quality of red blood cells in the bloodstream.

```{r, message=FALSE, warning=FALSE, eval=TRUE}
data <- data %>% 
  dplyr::select(sample,mt,sex,age,logmt, Batch,PC1,PC2,PC3,PC4,PC5, diagnosis,
                `White blood cell (leukocyte) count`,`Platelet count`,`Platelet crit`,
                `Mean platelet (thrombocyte) volume`,`Platelet distribution width`,`Lymphocyte count`,
                `Monocyte count`,`Neutrophill count`,`Eosinophill count`,
                `Lymphocyte percentage`,`Monocyte percentage`,`Neutrophill percentage`,
                `Eosinophill percentage`
                )

mtDNA_ukbb_summary_PD_blood_correction <- data

```


<span style="color:blue">The associations of various blood compositions with PD diagnosis</span>

```{r, message=FALSE, warning=FALSE, eval=TRUE}
data <- mtDNA_ukbb_summary_PD_blood_correction
data <- as.data.frame(data)
data$diagnosis <- factor(data$diagnosis, level = c("Control", "Case"))
data$sex <- factor(data$sex, level = c("Female", "Male"))
data$Batch <- factor(data$Batch, level = c("1", "2"))
data <- data %>% 
  na.omit()
data <- data %>% 
  mutate(NLR=`Neutrophill percentage`/`Lymphocyte percentage`)
measure <- colnames(data)[13:ncol(data)]
result <- list()
for (i in 1:length(measure)) {
res <- summary(lm.dx <- rlm(formula = data[,measure[i]] ~ diagnosis + age + sex + Batch + PC1 + PC2 + PC3 + PC4 + PC5, 
                            data = data))$coefficients[2,] %>% unlist %>% matrix(nrow=1) %>% as.data.frame

res$p.value <- f.robftest(lm.dx, var = "diagnosisCase")$p.value
res$n.samp <- nrow(data)
colnames(res)[1:4] <- c("Estimate","Std. Error","t value","Pr(>|t|)")
res <- res %>% 
  mutate(Measure=measure[i])
result[[i]] <- res
}

rlm_result <- do.call(rbind,result)
rlm_result$p.adj <- p.adjust(rlm_result$`Pr(>|t|)`, method="fdr")

rlm_result_end <- data.frame(
  Blood_measurement=rlm_result$Measure,
  Estimate= round(rlm_result$Estimate,3),
  `Std. Error`= round(rlm_result$`Std. Error`,3),
  `t value` = round(rlm_result$`t value`,2),
  `Pr(>|t|)` = ifelse(rlm_result$`Pr(>|t|)` < 0.0001,"p<0.0001",round(rlm_result$`Pr(>|t|)`,4)),
  p.adj=ifelse(rlm_result$p.adj < 0.0001,"p<0.0001",round(rlm_result$p.adj,4))
)

colnames(rlm_result_end) <- c("Blood_measurement","Estimate","Std. Error","t value","Pr(>|t|)", "P.adj")
datatable(rlm_result_end,
  extensions = "Buttons",
  rownames = FALSE,
  options = list(
    pageLength = 15,
    dom = "Blfrtip",
    buttons = c("copy", "csv", "excel")
  ),
  escape = F
)

```

### Association test of raw and adjusted blood mtDNA-CN with PD risk 

```{r, message=FALSE, warning=FALSE, eval=TRUE}
#Blood count-based measure
data <- mtDNA_ukbb_summary_PD_blood_correction
#Select cell types 
data <- data %>% 
  dplyr::select(sample,mt,sex,age,logmt, Batch,PC1,PC2,PC3,PC4,PC5,diagnosis,
                `White blood cell (leukocyte) count`,`Platelet count`,`Platelet crit`,
                `Mean platelet (thrombocyte) volume`,`Platelet distribution width`,
                `Lymphocyte percentage`,`Monocyte percentage`,`Neutrophill percentage`,`Eosinophill percentage`
                )
mt_data <- data %>% 
  na.omit()
### model selection
# full model
m0 <- lm(logmt ~ ., data = mt_data[,-c(1,2,3,4,6:12)])
summary(m0)
m1 <- step(m0, direction = "both")
summary(m1)
```



<span style="color:blue">**Adjust the logmt (correcting for blood composition)**</span>

```{r, message=FALSE, warning=FALSE, eval=TRUE}
# cell composition correction
mt.adj <- round(residuals(m1),4)
adj.mt <- cbind(mt_data, mt.adj)
adj.mt <- adj.mt %>% 
  dplyr::select(sample,mt,logmt,mt.adj,Batch,sex,age,PC1,  PC2, PC3, PC4, PC5, diagnosis)
adj.mt %>% 
  as.tibble()
mtDNA_ukbb_summary_PD_blood_correction_adj.mt <- adj.mt
```

<br>

<hr>


<span style="color:blue">mtDNA-CN (raw) and PD diagnosis</span>

```{r, message=FALSE, warning=FALSE, eval=TRUE}
# association test
data <- mtDNA_ukbb_summary_PD_blood_correction_adj.mt
data$diagnosis <- factor(data$diagnosis, level = c("Control", "Case"))
data$sex <- factor(data$sex, level = c("Female", "Male"))
data$Batch <- factor(data$Batch, level = c("1", "2"))
summary(data)
##### diagnosis
test = data
res <- summary(lm.dx <- rlm(formula = logmt ~ diagnosis + sex + age + Batch + PC1 + PC2 + PC3 + PC4 + PC5, 
                            data = test))$coefficients[2,] %>% unlist %>% matrix(nrow=1) %>% as.data.frame
res$p.value <- f.robftest(lm.dx, var = "diagnosisCase")$p.value
res$n.samp <- nrow(test)
res$Control <- paste0(round(mean(test$mt[test$diagnosis == "Control"]),2),
                      "(",round(sd(test$mt[test$diagnosis == "Control"]),2),")")
res$Case <- paste0(round(mean(test$mt[test$diagnosis == "Case"]),2),
                   "(",round(sd(test$mt[test$diagnosis == "Case"]),2),")")
colnames(res)[1:4] <- c("Estimate","Std. Error","t value","Pr(>|t|)")
res$cluster <- c(2)
res$study <- c("UKBB")
res
res_raw <- res %>% 
  mutate(type="logmtRaw")
```



<br>

<span style="color:blue">mtDNA-CN (adjusted) and PD diagnosis</span>


```{r, message=FALSE, warning=FALSE, eval=TRUE}
# association test
data <- mtDNA_ukbb_summary_PD_blood_correction_adj.mt
# association test
data$diagnosis <- factor(data$diagnosis, level = c("Control", "Case"))
data$sex <- factor(data$sex, level = c("Female", "Male"))
data$Batch <- factor(data$Batch, level = c("1", "2"))
summary(data)
##### diagnosis
test = data
res <- summary(lm.dx <- rlm(formula = mt.adj ~ diagnosis + sex + age + Batch + PC1 + PC2 + PC3 + PC4 + PC5, 
                            data = test))$coefficients[2,] %>% unlist %>% matrix(nrow=1) %>% as.data.frame
res$p.value <- f.robftest(lm.dx, var = "diagnosisCase")$p.value
res$n.samp <- nrow(test)
res$Control <- paste0(round(mean(test$mt[test$diagnosis == "Control"]),2),
                      "(",round(sd(test$mt[test$diagnosis == "Control"]),2),")")
res$Case <- paste0(round(mean(test$mt[test$diagnosis == "Case"]),2),
                   "(",round(sd(test$mt[test$diagnosis == "Case"]),2),")")
colnames(res)[1:4] <- c("Estimate","Std. Error","t value","Pr(>|t|)")
res$cluster <- c(2)
res$study <- c("UKBB")
res
res_adj <- res %>% 
  mutate(type="logmtAdjusted")
res_all <- do.call(rbind,list(res_raw,res_adj))
res_all <- res_all %>% 
  mutate(SelectVariable=c("", "White blood cell (leukocyte) count; Platelet count; Platelet crit; Mean platelet (thrombocyte) volume; Platelet distribution width; Lymphocyte percentage; Monocyte percentage; Neutrophill percentage; Eosinophill percentage"))
```

<br>


### Association test results of raw and adjusted blood mtDNA-CN with PD risk in the UKB cohort

```{r, message=FALSE, warning=FALSE, eval=TRUE}
#Test the associations between mtDNA-CN (raw vs adjusted) and PD diagnosis
#mtDNA-CN raw
# association test
data <- mtDNA_ukbb_summary_PD_blood_correction_adj.mt
data$diagnosis <- factor(data$diagnosis, level = c("Control", "Case"))
data$sex <- factor(data$sex, level = c("Female", "Male"))
data$Batch <- factor(data$Batch, level = c("1", "2"))
summary(data)
#age
test = data
res <- summary(lm.dx <- rlm(formula = logmt ~ diagnosis + sex + age + Batch + PC1 + PC2 + PC3 + PC4 + PC5, 
                            data = test))$coefficients

#confindence level
confint_diagnosisCase <- confint.default(object = lm.dx, parm = "diagnosisCase", level = 0.95)

res <- data.frame(
  var=rownames(res),
  res
) %>% 
 filter(var %in% c("diagnosisCase","sexMale","age","Batch2"))
res$p.value[1] <- f.robftest(lm.dx, var = "diagnosisCase")$p.value
res$p.value[2] <- f.robftest(lm.dx, var = "sexMale")$p.value
res$p.value[3] <- f.robftest(lm.dx, var = "age")$p.value
res$p.value[4] <- f.robftest(lm.dx, var = "Batch2")$p.value
res$n.samp <- nrow(test)
colnames(res)[1:6] <- c("Variable","Estimate","Std. Error","t value","Pr(>|t|)","Nsample")
res_raw <- res
res_raw <- data.frame(
  Type="mtCNraw",
  res_raw
)
colnames(res_raw)[1:7] <- c("Type","Variable","Estimate","Std. Error","t value","Pr(>|t|)","Nsample")

#mtDNA-CN adjusted
# association test
data <- mtDNA_ukbb_summary_PD_blood_correction_adj.mt
data$diagnosis <- factor(data$diagnosis, level = c("Control", "Case"))
data$sex <- factor(data$sex, level = c("Female", "Male"))
data$Batch <- factor(data$Batch, level = c("1", "2"))
#summary(data)
#age
test = data
res <- summary(lm.dx <- rlm(formula = mt.adj ~ diagnosis + sex + age + Batch + PC1 + PC2 + PC3 + PC4 + PC5, 
                            data = test))$coefficients

#confindence level
confint_diagnosisCase <- confint.default(object = lm.dx, parm = "diagnosisCase", level = 0.95)

res <- data.frame(
  var=rownames(res),
  res
) %>% 
 filter(var %in% c("diagnosisCase","sexMale","age","Batch2"))
res$p.value[1] <- f.robftest(lm.dx, var = "diagnosisCase")$p.value
res$p.value[2] <- f.robftest(lm.dx, var = "sexMale")$p.value
res$p.value[3] <- f.robftest(lm.dx, var = "age")$p.value
res$p.value[4] <- f.robftest(lm.dx, var = "Batch2")$p.value
res$n.samp <- nrow(test)
colnames(res)[1:6] <- c("Variable","Estimate","Std. Error","t value","Pr(>|t|)","Nsample")
res_adj <- res
res_adj <- data.frame(
  Type="mtCNadj",
  res_adj
)
colnames(res_adj)[1:7] <- c("Type","Variable","Estimate","Std. Error","t value","Pr(>|t|)","Nsample")

res_all <- do.call(rbind,list(res_raw,res_adj))
res_all_end <- data.frame(
  Type=res_all$Type,
  Variable=res_all$Variable,
  Estimate= round(res_all$Estimate,3),
  `Std. Error`= round(res_all$`Std. Error`,3),
  `t value` = round(res_all$`t value`,2),
  `Pr(>|t|)` = ifelse(res_all$`Pr(>|t|)` < 0.0001,"p<0.0001",round(res_all$`Pr(>|t|)`,3))
)

datatable(res_all_end,
          rownames = FALSE,
          extensions = 'Buttons',
          options = list(pageLength = 10,
                          dom = 'Blfrtip',
                          buttons = c('copy', 'csv', 'excel')),
          escape = F)
```
