---
title: "UKB mtDNA-CN Analysis"
author: "Jiru Han"
output:
  html_document:
    code_folding: "hide"
  df_print: paged
---



```{r setup, include=FALSE} 
library(tidyverse)
library(magrittr)
library(data.table)
library(pander)
library(plotly)
library(KRIS)
library(MASS)
library(sfsmisc)
library(DT)
library(ggpubr)
library(ggstatsplot)
select <- dplyr::select

# formatting functions
sensible_names <- function(x){
  colnames(x) %<>%
    str_replace_all('[ -//]+', '.') %>%
    str_to_lower()
  x
}

knitr::opts_chunk$set(tidy="styler")
```


<span style="color:blue">**Check Batch effects (cases)**</span>

```{r,message=FALSE, warning=FALSE, eval=TRUE}
ukbdata <- fread("/vast/projects/bahlo_ukbiobank/app36610_HDA/mitochondria/20240423_mtDNA_ukbb_batch1_batch2_summary.csv")
batch <- fread("/vast/projects/bahlo_ukbiobank/app36610_HDA/mitochondria/20240423_UKBB_eid_sequencing_batch.tsv")
ukbdata <- ukbdata %>% 
  full_join(batch,by=c("sample"="eid"))
#Withdrawn participants
withdrawn <- fread("/vast/projects/bahlo_ukbiobank/app36610_HDA/w36610_2023-04-25.csv")
ukbdata <- ukbdata %>% 
  filter(!(sample %in% withdrawn$V1))
#485443
all <- ukbdata %>% 
  filter(Parkinsons=="Y") %>% 
  dplyr::select(sample,batch, genetic_sex, mt,Blood_Collection_Age,logmt)
colnames(all)[2] <- "Data"
all$Data <- as.character(all$Data)
all %>% 
  pivot_longer(cols = mt:logmt,
               names_to = "Type",
               values_to = "Value") %>% 
  filter(Type=="logmt") %>% 
  ggplot(aes(x=Value,fill=Data))+
  geom_density()+
  labs(x="logmt")

all %>% 
  dplyr::select(sample,Data,genetic_sex, mt,Blood_Collection_Age,logmt) %>% 
  pivot_longer(cols = mt:logmt,
               names_to = "Type",
               values_to = "Value") %>% 
  filter(Type=="logmt") %>% 
  ggplot(aes(x=Data,y=Value,fill=Data))+
  geom_boxplot()+
  labs(y="logmt")


old <- all %>% 
  filter(Data=="1")
new <- all %>% 
  filter(Data=="2")
t.test(old$logmt, y = new$logmt,
       alternative = "two.sided")
```


<br>

<span style="color:blue">UKBBdata table (combine boold cell composition & remove NA)</span>

```{r,message=FALSE, warning=FALSE, eval=TRUE}
# load mtDNA-CN results
ukbdata <- fread("/vast/projects/bahlo_ukbiobank/app36610_HDA/mitochondria/20240423_mtDNA_ukbb_batch1_batch2_summary.csv")
batch <- fread("/vast/projects/bahlo_ukbiobank/app36610_HDA/mitochondria/20240423_UKBB_eid_sequencing_batch.tsv")
ukbdata <- ukbdata %>% 
  full_join(batch,by=c("sample"="eid"))
#Withdrawn participants
withdrawn <- fread("/vast/projects/bahlo_ukbiobank/app36610_HDA/w36610_2023-04-25.csv")
ukbdata <- ukbdata %>% 
  filter(!(sample %in% withdrawn$V1))
#485443

ukbdata <- ukbdata %>% 
  dplyr::select(sample,mt,genetic_sex,Blood_Collection_Age,logmt,batch,PC1,PC2,PC3,PC4,PC5,Parkinsons)
colnames(ukbdata)[6] <- "Batch"

ukbdata <- ukbdata %>% 
  dplyr::select(sample,mt,genetic_sex,Blood_Collection_Age,logmt,Batch,PC1,PC2,PC3,PC4,PC5,Parkinsons) %>% 
  mutate(PD_diagnosis = ifelse(Parkinsons=="Y","Case","Control")) %>% 
  dplyr::select(-Parkinsons) %>% 
  na.omit()
colnames(ukbdata) <- c("sample", "mt", "sex", "age","logmt", "Batch", "PC1","PC2", "PC3","PC4", "PC5", "diagnosis")
#boold cell composition
data <- fread("/vast/projects/bahlo_ukbiobank/app36610_HDA/extracts/20240118_673741_ukb_blood.csv.gz")
data <- data %>% 
  dplyr::select(eid,matches("0-0.0")) %>% 
  dplyr::select(eid,where(is.numeric))
data <- data %>%   
  pivot_longer(!eid,
               names_to = "FieldIDNew",
               values_to = "Value")
dic <- fread("/vast/projects/bahlo_ukbiobank/app36610_HDA/extracts/20240219_Data_Dictionary_Showcase.csv")
dic <- dic %>% 
  dplyr::select(FieldID,Field)
dic <- dic %>% 
  mutate(FieldIDNew=paste0(FieldID,"-0.0"))
data <- data %>% 
  left_join(dic)
data <- data %>% 
  dplyr::select(eid,Field,Value) %>% 
  pivot_wider(names_from = Field,
              values_from = Value)
ukbdata <- ukbdata %>% 
  left_join(data,by=c("sample"="eid"))
ukbdata %>% 
   as.tibble()
mtDNA_ukbb_summary_PD_blood <- ukbdata
```



<br>

### 1. Check log(mtDNA-CN) distribution using histogram.

<span style="color:blue">**logmt**</span>

```{r,message=FALSE, warning=FALSE, eval=TRUE}
ukbdata <- mtDNA_ukbb_summary_PD_blood
range(ukbdata$logmt)
hist(ukbdata$logmt)
```


<br>

<hr>

### 2. Stepwise model section with both direction:

logmt ~ blood cell variables

adjusted mt: residual of logmt ~ selected variables
 
### Adjust the boold cell composition

<span style="color:blue">**UKBB boold cell measurements**</span>


```{r, message=FALSE, warning=FALSE, eval=TRUE}
blood_measure <- read.csv("Blood_Measurements.csv")
blood_measure$UKBBLink <- paste0('<a href="',blood_measure$UKBBLink,'">', 'UKBB_', blood_measure$FieldID ,"</a>")
datatable(blood_measure,
          extensions = 'Buttons',
          options = list(pageLength = 10,
                          dom = 'Blfrtip',
                          buttons = c('copy', 'csv', 'excel')),
          escape = F)
```

<br>

<span style="color:blue">**UKBB boold cell composition data**</span>

### Cell composition correction

<span style="color:blue">Remove high missing blood cell type > 1%</span>

```{r, message=FALSE, warning=FALSE, eval=TRUE}
#Remove high missing cell type
data <- mtDNA_ukbb_summary_PD_blood
data_miss <- data.frame(miss=colSums(is.na(data))/nrow(data)*100) %>% 
  filter(miss>1)
data_miss
data <- data %>% 
  dplyr::select(-rownames(data_miss))
```

<span style="color:blue">Blood cell type QC: select cell types percentage with proportion > 1%</span>

remove "Basophill percentage/count","Nucleated red blood cell percentage/count"

```{r, message=FALSE, warning=FALSE, eval=TRUE}
# QC: select cell types percentage with proportion > 1%
cell_percentage <- data %>% 
  dplyr::select(sample,matches("percentage"))
vars <- c(colnames(cell_percentage[,2:8])[colMeans(cell_percentage[,2:8], na.rm = TRUE) < 1])
# remove "Basophill percentage","Nucleated red blood cell percentage"
vars
#colMeans(cell_percentage[,2:8], na.rm = TRUE)
#Basophill percentage 0.56998017 
#Nucleated red blood cell percentage 0.03505115 
data <- data %>% 
  dplyr::select(-c(vars,`Basophill count`,`Nucleated red blood cell count`))
```



<span style="color:blue">**Remove all the RBC measurements**</span>

The blood measurements associated with Red Blood Cells (RBCs) are:
Red blood cell (erythrocyte) count
Haemoglobin concentration
Haematocrit percentage
Mean corpuscular volume
Mean corpuscular haemoglobin
Mean corpuscular haemoglobin concentration
Red blood cell (erythrocyte) distribution width
Nucleated red blood cell count
Nucleated red blood cell percentage
Mean reticulocyte volume
Mean sphered cell volume
These measurements provide valuable information about the quantity, size, and quality of red blood cells in the bloodstream.

```{r, message=FALSE, warning=FALSE, eval=TRUE}
data <- data %>% 
  dplyr::select(sample,mt,sex,age,logmt, Batch,PC1,PC2,PC3,PC4,PC5, diagnosis,
                `White blood cell (leukocyte) count`,`Platelet count`,`Platelet crit`,
                `Mean platelet (thrombocyte) volume`,`Platelet distribution width`,`Lymphocyte count`,
                `Monocyte count`,`Neutrophill count`,`Eosinophill count`,
                `Lymphocyte percentage`,`Monocyte percentage`,`Neutrophill percentage`,
                `Eosinophill percentage`
                )

mtDNA_ukbb_summary_PD_blood_correction <- data

```



<span style="color:blue">**Boold cell composition correlation matrix plot without RBC measurements (after remove high missing rate > 1% & cell types percentage with proportion < 1%)**</span>


```{r, message=FALSE, warning=FALSE, eval=TRUE,fig.width=15,fig.height=15}
data <- mtDNA_ukbb_summary_PD_blood_correction
ggcorrmat(data[,-c(1,2,3,4,6:12)],colors   = c("#B2182B", "white", "#4D4D4D"),
          ggcorrplot.args = list(method = "square", outline.color = "black",lab_size = 3),
          number.cex = 0.1,
          matrix.type  = "lower")
```


<span style="color:blue">**PD_diagnosis & Boold cell composition**</span>

<span style="color:blue">blood var rlm ~ PD_diagnosis + age + sex + Batch + PC1-5</span>

blood var rlm ~ PD_diagnosis estimate



```{r, message=FALSE, warning=FALSE, eval=TRUE}
data <- mtDNA_ukbb_summary_PD_blood_correction
data <- as.data.frame(data)
data$diagnosis <- factor(data$diagnosis, level = c("Control", "Case"))
data$sex <- factor(data$sex, level = c("Female", "Male"))
data$Batch <- factor(data$Batch, level = c("1", "2"))
data <- data %>% 
  na.omit()
data <- data %>% 
  mutate(NLR=`Neutrophill percentage`/`Lymphocyte percentage`)
#431093
#Check relatedness
related_remove <- fread("/vast/projects/bahlo_ukbiobank/app36610_HDA/mitochondria/20240426_mwvc_pd_todrop.txt_MWVC.txt")
data <- data %>% 
  filter(!(sample %in% related_remove$V1))
#367322
measure <- colnames(data)[13:ncol(data)]
result <- list()
for (i in 1:length(measure)) {
res <- summary(lm.dx <- rlm(formula = data[,measure[i]] ~ diagnosis + age + sex + Batch + PC1 + PC2 + PC3 + PC4 + PC5, 
                            data = data))$coefficients[2,] %>% unlist %>% matrix(nrow=1) %>% as.data.frame

res$p.value <- f.robftest(lm.dx, var = "diagnosisCase")$p.value
res$n.samp <- nrow(data)
colnames(res)[1:4] <- c("Estimate","Std. Error","t value","Pr(>|t|)")
res <- res %>% 
  mutate(Measure=measure[i])
result[[i]] <- res
}

rlm_result <- do.call(rbind,result)
rlm_result$p.adj <- p.adjust(rlm_result$`Pr(>|t|)`, method="fdr")

rlm_result_end <- data.frame(
  Blood_measurement=rlm_result$Measure,
  Estimate= round(rlm_result$Estimate,3),
  `Std. Error`= round(rlm_result$`Std. Error`,3),
  `t value` = round(rlm_result$`t value`,2),
  `Pr(>|t|)` = ifelse(rlm_result$`Pr(>|t|)` < 0.0001,"p<0.0001",round(rlm_result$`Pr(>|t|)`,4)),
  p.adj=ifelse(rlm_result$p.adj < 0.0001,"p<0.0001",round(rlm_result$p.adj,4))
)


colnames(rlm_result_end) <- c("Blood_measurement","Estimate","Std. Error","t value","Pr(>|t|)", "P.adj")
#write.csv(rlm_result_end,"Blood_measurement_PDdiagnosis.csv",row.names = FALSE)

datatable(rlm_result_end,
  extensions = "Buttons",
  rownames = FALSE,
  options = list(
    pageLength = 15,
    dom = "Blfrtip",
    buttons = c("copy", "csv", "excel")
  ),
  escape = F
)

```


### Model selection


<span style="color:blue">**Blood percentage & count-based measure**</span>


Because some measurements do not have percentages, we will still use counts instead for these measurements


**Full model**

```{r, message=FALSE, warning=FALSE, eval=TRUE}
#Blood count-based measure
data <- mtDNA_ukbb_summary_PD_blood_correction
#Select cell types 
data <- data %>% 
  dplyr::select(sample,mt,sex,age,logmt, Batch,PC1,PC2,PC3,PC4,PC5,diagnosis,
                `White blood cell (leukocyte) count`,`Platelet count`,`Platelet crit`,
                `Mean platelet (thrombocyte) volume`,`Platelet distribution width`,
                #`Lymphocyte count`,`Monocyte count`,`Neutrophill count`,`Eosinophill count`
                `Lymphocyte percentage`,`Monocyte percentage`,`Neutrophill percentage`,`Eosinophill percentage`
                )
#Check relatedness
related_remove <- fread("/vast/projects/bahlo_ukbiobank/app36610_HDA/mitochondria/20240426_mwvc_pd_todrop.txt_MWVC.txt")
data <- data %>% 
  filter(!(sample %in% related_remove$V1))
mt_data <- data %>% 
  na.omit()
#367322     21
### model selection
# full model
m0 <- lm(logmt ~ ., data = mt_data[,-c(1,2,3,4,6:12)])
summary(m0)
```


<br>

```{r, message=FALSE, warning=FALSE, eval=TRUE,results='hide'}
m1 <- step(m0, direction = "both")
```



**Step-wise regression**

```{r, message=FALSE, warning=FALSE, eval=TRUE}
summary(m1) #R2 = 19%
```



### Adjust the logmt

```{r, message=FALSE, warning=FALSE, eval=TRUE}
# cell composition correction
mt.adj <- round(residuals(m1),4)
adj.mt <- cbind(mt_data, mt.adj)
adj.mt <- adj.mt %>% 
  dplyr::select(sample,mt,logmt,mt.adj,Batch,sex,age,PC1,  PC2, PC3, PC4, PC5, diagnosis)
adj.mt %>% 
  as.tibble()

mtDNA_ukbb_summary_PD_blood_correction_adj.mt <- adj.mt
```

<br>

<hr>


# Step 3
Adjusted mt ~ PD diagnosis + age + sex + Batch + PC1-5


### Test the associations between mtDNA-CN (raw vs adjusted) and PD diagnosis


<span style="color:blue">mtDNA-CN (raw) and PD diagnosis</span>

```{r, message=FALSE, warning=FALSE, eval=TRUE}
# association test
data <- mtDNA_ukbb_summary_PD_blood_correction_adj.mt
# association test
data$diagnosis <- factor(data$diagnosis, level = c("Control", "Case"))
data$sex <- factor(data$sex, level = c("Female", "Male"))
data$Batch <- factor(data$Batch, level = c("1", "2"))
summary(data)
##### diagnosis
test = data
res <- summary(lm.dx <- rlm(formula = logmt ~ diagnosis + sex + age + Batch + PC1 + PC2 + PC3 + PC4 + PC5, 
                            data = test))$coefficients[2,] %>% unlist %>% matrix(nrow=1) %>% as.data.frame
res$p.value <- f.robftest(lm.dx, var = "diagnosisCase")$p.value
res$n.samp <- nrow(test)
res$Control <- paste0(round(mean(test$mt[test$diagnosis == "Control"]),2),
                      "(",round(sd(test$mt[test$diagnosis == "Control"]),2),")")
res$Case <- paste0(round(mean(test$mt[test$diagnosis == "Case"]),2),
                   "(",round(sd(test$mt[test$diagnosis == "Case"]),2),")")
colnames(res)[1:4] <- c("Estimate","Std. Error","t value","Pr(>|t|)")
res$cluster <- c(2)
res$study <- c("UKBB")
res
res_raw <- res %>% 
  mutate(type="logmtRaw")
```



<br>

<span style="color:blue">mtDNA-CN (adjusted) and PD diagnosis</span>


```{r, message=FALSE, warning=FALSE, eval=TRUE}
# association test
data <- mtDNA_ukbb_summary_PD_blood_correction_adj.mt
# association test
data$diagnosis <- factor(data$diagnosis, level = c("Control", "Case"))
data$sex <- factor(data$sex, level = c("Female", "Male"))
data$Batch <- factor(data$Batch, level = c("1", "2"))
summary(data)
##### diagnosis
test = data
res <- summary(lm.dx <- rlm(formula = mt.adj ~ diagnosis + sex + age + Batch + PC1 + PC2 + PC3 + PC4 + PC5, 
                            data = test))$coefficients[2,] %>% unlist %>% matrix(nrow=1) %>% as.data.frame
res$p.value <- f.robftest(lm.dx, var = "diagnosisCase")$p.value
res$n.samp <- nrow(test)
res$Control <- paste0(round(mean(test$mt[test$diagnosis == "Control"]),2),
                      "(",round(sd(test$mt[test$diagnosis == "Control"]),2),")")
res$Case <- paste0(round(mean(test$mt[test$diagnosis == "Case"]),2),
                   "(",round(sd(test$mt[test$diagnosis == "Case"]),2),")")
colnames(res)[1:4] <- c("Estimate","Std. Error","t value","Pr(>|t|)")
res$cluster <- c(2)
res$study <- c("UKBB")
res
res_adj <- res %>% 
  mutate(type="logmtAdjusted")
```


<br>

<hr>



# Step 4

Make a table to compare between raw and adjusted mt


```{r, message=FALSE, warning=FALSE, eval=TRUE}
res_all <- do.call(rbind,list(res_raw,res_adj))
res_all <- res_all %>% 
  mutate(SelectVariable=c("", "White blood cell (leukocyte) count; Platelet count; Platelet crit; Mean platelet (thrombocyte) volume; Platelet distribution width; Lymphocyte percentage; Monocyte percentage; Neutrophill percentage; Eosinophill percentage"))
res_all
```











<br>


<span style="color:blue">Scatter plot age vs logmt & adjusted logmt</span>

```{r,warning=FALSE,message=FALSE,echo=FALSE}
theme_Publication <- function(base_size=14, base_family="helvetica") {
      library(grid)
      library(ggthemes)
      (theme_foundation(base_size=base_size, base_family=base_family)
       + theme(plot.title = element_text(face = "bold",
                                         size = rel(1.2), hjust = 0.5),
               text = element_text(),
               panel.background = element_rect(colour = NA),
               plot.background = element_rect(colour = NA),
               panel.border = element_rect(colour = NA),
               axis.title = element_text(face = "bold",size = rel(1)),
               axis.title.y = element_text(angle=90,vjust =2),
               axis.title.x = element_text(vjust = -0.2),
               axis.text = element_text(), 
               axis.line = element_line(colour="black"),
               axis.ticks = element_line(),
               panel.grid.major = element_line(colour="#f0f0f0"),
               panel.grid.minor = element_blank(),
               legend.key = element_rect(colour = NA),
               legend.position = "bottom",
               legend.direction = "horizontal",
               legend.key.size= unit(0.2, "cm"),
               legend.margin = unit(0, "cm"),
               legend.title = element_text(face="italic"),
               plot.margin=unit(c(10,5,5,5),"mm"),
               strip.background=element_rect(colour="#f0f0f0",fill="#f0f0f0"),
               strip.text = element_text(face="bold")
          ))
      
}

```




```{r,message=FALSE, warning=FALSE, eval=TRUE}
#367322
data <- mtDNA_ukbb_summary_PD_blood_correction_adj.mt
scatter_plot <- data %>% 
  dplyr::select(sample,sex,age,diagnosis,Batch,logmt,mt.adj) %>% 
  pivot_longer(cols = logmt:mt.adj,
               names_to = "logmt",
               values_to = "Value")


scatter_plot$logmt[scatter_plot$logmt =="logmt"] <- "mtCNraw"
scatter_plot$logmt[scatter_plot$logmt =="mt.adj"] <- "mtCNadj"

scatter_plot$logmt <- factor(scatter_plot$logmt,levels = c("mtCNraw","mtCNadj"))
colnames(scatter_plot)[2] <- "Sex"


#mtCNraw

p1 <- scatter_plot %>% 
  filter(logmt=="mtCNraw") %>% 
  ggplot(aes(x=age,y=Value,color=Sex))+
  geom_point(alpha=0.1)+
  geom_smooth(method = "lm",size = 2)+
  stat_cor()+
  #facet_grid(logmt~.,space = "free",scales = "free")+
  ggsci::scale_color_npg()+
  #theme_bw()+
  #ggthemes::theme_calc()+
  theme_Publication()+
  labs(x="Age",y="log-transformed mtDNA copy number")


p2 <- scatter_plot %>% 
  filter(logmt=="mtCNadj") %>% 
  ggplot(aes(x=age,y=Value,color=Sex))+
  geom_point(alpha=0.1)+
  geom_smooth(method = "lm",size = 2)+
  stat_cor()+
  #facet_grid(logmt~.,space = "free",scales = "free")+
  ggsci::scale_color_npg()+
  #theme_bw()+
  #ggthemes::theme_calc()+
  theme_Publication()+
  labs(x="Age",y="Adjusted mtDNA copy number")
#(Blood cell composition correction)

ggpubr::ggarrange(p1,p2,nrow=1,common.legend = TRUE,legend="bottom")
#1500 700
```


<br>


### Summary table logmt & mt.adj ~ sex, age, Batch, diagnosis, PC1-5

```{r, message=FALSE, warning=FALSE, eval=TRUE}
#Test the associations between mtDNA-CN (raw vs adjusted) and PD diagnosis
#mtDNA-CN raw
# association test
data <- mtDNA_ukbb_summary_PD_blood_correction_adj.mt
# association test
data$diagnosis <- factor(data$diagnosis, level = c("Control", "Case"))
data$sex <- factor(data$sex, level = c("Female", "Male"))
data$Batch <- factor(data$Batch, level = c("1", "2"))
summary(data)
#age
test = data
res <- summary(lm.dx <- rlm(formula = logmt ~ diagnosis + sex + age + Batch + PC1 + PC2 + PC3 + PC4 + PC5, 
                            data = test))$coefficients

#confindence level
confint.default(object = lm.dx, parm = "diagnosisCase", level = 0.95)

res <- data.frame(
  var=rownames(res),
  res
) %>% 
 filter(var %in% c("diagnosisCase","sexMale","age","Batch2"))
res$p.value[1] <- f.robftest(lm.dx, var = "diagnosisCase")$p.value
res$p.value[2] <- f.robftest(lm.dx, var = "sexMale")$p.value
res$p.value[3] <- f.robftest(lm.dx, var = "age")$p.value
res$p.value[4] <- f.robftest(lm.dx, var = "Batch2")$p.value
res$n.samp <- nrow(test)
colnames(res)[1:6] <- c("Variable","Estimate","Std. Error","t value","Pr(>|t|)","Nsample")
res_raw <- res
res_raw <- data.frame(
  Type="mtCNraw",
  res_raw
)
colnames(res_raw)[1:7] <- c("Type","Variable","Estimate","Std. Error","t value","Pr(>|t|)","Nsample")




#mtDNA-CN adjusted
# association test
data <- mtDNA_ukbb_summary_PD_blood_correction_adj.mt
# association test
data$diagnosis <- factor(data$diagnosis, level = c("Control", "Case"))
data$sex <- factor(data$sex, level = c("Female", "Male"))
data$Batch <- factor(data$Batch, level = c("1", "2"))
#summary(data)
#age
test = data
res <- summary(lm.dx <- rlm(formula = mt.adj ~ diagnosis + sex + age + Batch + PC1 + PC2 + PC3 + PC4 + PC5, 
                            data = test))$coefficients

#confindence level
confint.default(object = lm.dx, parm = "diagnosisCase", level = 0.95)

res <- data.frame(
  var=rownames(res),
  res
) %>% 
 filter(var %in% c("diagnosisCase","sexMale","age","Batch2"))
res$p.value[1] <- f.robftest(lm.dx, var = "diagnosisCase")$p.value
res$p.value[2] <- f.robftest(lm.dx, var = "sexMale")$p.value
res$p.value[3] <- f.robftest(lm.dx, var = "age")$p.value
res$p.value[4] <- f.robftest(lm.dx, var = "Batch2")$p.value
res$n.samp <- nrow(test)
colnames(res)[1:6] <- c("Variable","Estimate","Std. Error","t value","Pr(>|t|)","Nsample")
res_adj <- res
res_adj <- data.frame(
  Type="mtCNadj",
  res_adj
)
colnames(res_adj)[1:7] <- c("Type","Variable","Estimate","Std. Error","t value","Pr(>|t|)","Nsample")

res_all <- do.call(rbind,list(res_raw,res_adj))
res_all_end <- data.frame(
  Type=res_all$Type,
  Variable=res_all$Variable,
  Estimate= round(res_all$Estimate,3),
  `Std. Error`= round(res_all$`Std. Error`,3),
  `t value` = round(res_all$`t value`,2),
  `Pr(>|t|)` = ifelse(res_all$`Pr(>|t|)` < 0.0001,"p<0.0001",round(res_all$`Pr(>|t|)`,3))
)

datatable(res_all_end,
          rownames = FALSE,
          extensions = 'Buttons',
          options = list(pageLength = 10,
                          dom = 'Blfrtip',
                          buttons = c('copy', 'csv', 'excel')),
          escape = F)
```




